{\fontfamily{ptm}\selectfont A Proximal Point framework of Nesterov type accelerated gradient}

### 2024-08-07 Notebook LaTEX/Momentum Accelerations/main.tex
>w<, work home sync

        \begin{theorem}[Constraints on the PPM stepsize sequence]
            \label{thm:ag_generic_stepsize_constrants}
            \;\\
            Algorithm formulated in 
            \hyperref[def:ag_prox_grad_generic]{definition \ref*{def:ag_prox_grad_generic}}
            converges if the stepsize sequence $\eta_i, \tilde \eta_{i}$ satisfies
            \begin{align*}
                \begin{cases}
                    \tilde \eta_{t + 1} (\tilde \eta_{t + 1} - L^{-1})
                    - L^{-1} \sum_{i= 1}^{t}\tilde \eta_i 
                    = 
                    \epsilon_{t + 1} \tilde \eta_{t + 1}
                    & \forall t \in \mathbb N, 
                    \\
                    L \eta_t \tilde \eta_{t + 1} = \sum_{i=1}^{t}\tilde \eta_i 
                    & \forall t \in \mathbb N. 
                \end{cases}
            \end{align*}
            Here $\epsilon_{t + 1} = \tilde \eta_{t + 1} - \eta_t - L^{-1}$ is a sequence made characterizing the error between $\tilde \eta_{t + 1}, \eta_t - L^{-1}$. 
            Then the convergence rate can be formulated by
            \begin{align*}
                \Phi_{t + 1} - \Phi_t =
                \Upsilon_{1, t + 1}^\text{AG} + 
                \sigma_t\Upsilon_{1, t + 1}^{\text{AG}} 
                &\le \epsilon_{t + 1}\tilde\eta_{t + 1} \Vert \mathcal G_L(y_t)\Vert^2 \le \delta_{t + 1}.
            \end{align*}
            When $\eta_t - 0 \;\forall t\in \mathbb N$, the sequence $a_t:= a_t = 1 + L \eta_t = \tilde \eta_{t + 1} - \epsilon_{t + 1}$, the relation simplies to: 
            \begin{align*}
                \begin{cases}
                    a_{t + 1} &= (1/2)\left(
                    1 + \sqrt{1 + 4 c_{t + 1}}
                    \right), 
                    \\
                    c_{t + 1} &= a_t^2.     
                \end{cases}
            \end{align*}
            As a consequence: 
            \begin{enumerate}
                \item It makes $\delta_{t} = 0$ in the Lypunov analysis; 
                \item it simplifies the algorithm to
                \hyperref[def:ag_form_similar_tria_I]{definition\ref*{def:ag_form_similar_tria_I}}; 
                \item and it is the famous Nesterov momentum sequence. 
            \end{enumerate}

@@

        \begin{lemma}[The Lyapunov function from estimating sequence and function]
            \;\\
            Assume that $f$ bounded below with $f_* = \min_x f(x)$ and there exists minimizer $x_*$. 
            With $\phi_k$ being an estimating functions and convex, $\{x_k\}$ an estimating sequence for $\phi_k$, then it satisfies
            \begin{align*}
                f(x_k) - f_*
                &\le 
                \lambda_k(\phi_0(x_*) - f_*). 
            \end{align*}
            for all $k\in \mathbb N\cup \{0\}$. 
        \end{lemma}
        \begin{proof}
            Directly by the property of estimating sequence $x_k$, we have for all $k, x$
            \begin{align*}
                \phi_k(x) - f(x) \ge \phi_k^* - f(x) \ge f(x_k) - f(x), 
            \end{align*}
            so 
            \begin{align*}
                f(x_k) - f(x) \le \lambda_k (\phi_0(x) - f(x))
                \\
                \implies 
                f(x_k) - f_* \le \lambda_k (\phi_0(x_*) - f(x_*)). 
            \end{align*}
            This establishes the convergence rate of $f(x_k)$ assuming that $\lambda_k$ is sequence that converges to zero. 
        \end{proof}
    \subsection{Construction of the estimating sequence and function}
        The construction of the Nesterov estimating sequence and function is where most creativities are involved. 
        For convex $f$ that is differentiable with $L$-Lipschitz gradient and strong convexity constant $\mu \ge 0$, Nesterov suggested in his book \cite{nesterov_lectures_2018} that there exists sequence $y_k$, $\gamma_k$, where
        \begin{align*}
            l_f(x, y_k) &= 
            f(y_k) + \langle \nabla f(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_0 (x) &= 
            f(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k(x) + \alpha_k l_f(x; y_k)
            \\
            &= \phi_{k + 1}^2 + \frac{\gamma_k}{2}\Vert x - v_{k + 1}\Vert^2
        \end{align*}
        will verify as a Nesterov's estimating sequence and functions. 
        The derivation for the parameters: $\gamma_k, x_k, y_k$ for $k \in \mathbb N \cup \{0\}$, will produce the Nesterov acclerated gradient as presented in
        \hyperref[def:Nes2.2.7]{definition \ref*{def:Nes2.2.7}}
        if correctly conducted. 
        In this case, $\phi_k$ are simple quadratic centered $v_k$ with value $\phi_k$ with curvature $\gamma_k$. 
        \par
        However, the construction exhibits variations for different variants of the accelerated gradient algorithm. 
        A different construction of $\phi_k$ yield a different algorithm. 
        As shown in Nesterov's book \cite*[(2.2.63)]{nesterov_lectures_2018}, a projected gradient variant of the accelerated gradient method make use of the following estimating functions, sequence: 
        \begin{align*}
            \phi_0 (x) &= f(x_0) + \frac{\gamma_0}{2}\Vert x - x_0\Vert^2
            \\
            l_f(x; y_k) &= f(\mathcal T_ly_k) + 
            \frac{1}{2L}\Vert \mathcal G_L(y_k)\Vert^2 + 
            \langle \mathcal G_L(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k + 
            \alpha_k l_f(x; y_k), 
        \end{align*}
        Here, $G_L(x) = L(x - \mathcal T_L(x))$ and $T_L(x) = \Pi_Q(x - L^{-1}\nabla f(x))$, where $Q$ is a convex set. 
        It's equivalent to applying proximal gradient method with $h = f + g$ where $g = \delta_Q$.

>>>>

        \begin{theorem}[Constraints on the PPM stepsize sequence]
            \label{thm:ag_generic_stepsize_constrants}
            \;\\
            Algorithm formulated in 
            \hyperref[def:ag_prox_grad_generic]{definition \ref*{def:ag_prox_grad_generic}}
            converges if the stepsize sequence $\eta_i, \tilde \eta_{i}$ satisfies
            \begin{align*}
                \begin{cases}
                    \tilde \eta_{t + 1} (\tilde \eta_{t + 1} - L^{-1})
                    - L^{-1} \sum_{i= 1}^{t}\tilde \eta_i 
                    = 
                    \epsilon_{t + 1} \tilde \eta_{t + 1}
                    & \forall t \in \mathbb N, 
                    \\
                    L \eta_t \tilde \eta_{t + 1} = \sum_{i=1}^{t}\tilde \eta_i 
                    & \forall t \in \mathbb N. 
                \end{cases}
            \end{align*}
            Here $\epsilon_{t + 1} = \tilde \eta_{t + 1} - \eta_t - L^{-1}$ is a sequence made characterizing the error between $\tilde \eta_{t + 1}, \eta_t - L^{-1}$. 
            Then the convergence rate can be formulated by
            \begin{align*}
                \Phi_{t + 1} - \Phi_t =
                \Upsilon_{1, t + 1}^\text{AG} + 
                \sigma_t\Upsilon_{1, t + 1}^{\text{AG}} 
                &\le \epsilon_{t + 1}\tilde\eta_{t + 1} \Vert \mathcal G_L(y_t)\Vert^2 \le \delta_{t + 1}.
            \end{align*}
            When $\epsilon_t = 0 \;\forall t\in \mathbb N$, the sequence $a_t:= a_t = 1 + L \eta_t = \tilde \eta_{t + 1} - \epsilon_{t + 1}$, the relation simplies to: 
            \begin{align*}
                \begin{cases}
                    a_{t + 1} = (1/2)\left(
                    1 + \sqrt{1 + 4 c_{t + 1}}
                    \right), 
                    \\
                    c_{t + 1} = a_t^2.     
                \end{cases}
            \end{align*}
            As a consequence: 
            \begin{enumerate}
                \item It makes $\delta_{t} = 0$ in the Lypunov analysis; 
                \item it simplifies the algorithm to
                \hyperref[def:ag_form_similar_tria_I]{definition\ref*{def:ag_form_similar_tria_I}}; 
                \item and it is the famous Nesterov momentum sequence. 
            \end{enumerate}

@@

        \begin{lemma}[The Lyapunov function from estimating sequence and function]
            \;\\
            Assume that $f$ bounded below with $f_* = \min_x f(x)$ and there exists minimizer $x_*$. 
            With $\phi_k$ being an estimating functions and convex, $\{x_k\}$ an estimating sequence for $\phi_k$, then it satisfies
            \begin{align*}
                f(x_k) - f_*
                &\le 
                \lambda_k(\phi_0(x_*) - f_*). 
            \end{align*}
            for all $k\in \mathbb N\cup \{0\}$. 
        \end{lemma}
        \begin{proof}
            Directly by the property of estimating sequence $x_k$, we have for all $k, x$
            \begin{align*}
                \phi_k(x) - f(x) \ge \phi_k^* - f(x) \ge f(x_k) - f(x), 
            \end{align*}
            so 
            \begin{align*}
                f(x_k) - f(x) \le \lambda_k (\phi_0(x) - f(x))
                \\
                \implies 
                f(x_k) - f_* \le \lambda_k (\phi_0(x_*) - f(x_*)). 
            \end{align*}
            This establishes the convergence rate of $f(x_k)$ assuming that $\lambda_k$ is sequence that converges to zero. 
        \end{proof}
    \subsection{Construction of the estimating sequence and function}
        The construction of the Nesterov estimating sequence and function is where most creativities are involved. 
        For convex $f$ that is differentiable with $L$-Lipschitz gradient and strong convexity constant $\mu \ge 0$, Nesterov suggested in his book \cite{nesterov_lectures_2018} that there exists sequence $y_k$, $\gamma_k$, where
        \begin{align*}
            l_f(x, y_k) &= 
            f(y_k) + \langle \nabla f(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_0 (x) &= 
            f(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k(x) + \alpha_k l_f(x; y_k)
            \\
            &= \phi_{k + 1}^2 + \frac{\gamma_k}{2}\Vert x - v_{k + 1}\Vert^2
        \end{align*}
        will verify as a Nesterov's estimating sequence and functions. 
        The derivation for the parameters: $\gamma_k, x_k, y_k$ for $k \in \mathbb N \cup \{0\}$, will produce the Nesterov acclerated gradient as presented in
        \hyperref[def:Nes2.2.7]{definition \ref*{def:Nes2.2.7}}
        if correctly conducted. 
        In this case, $\phi_k$ are simple quadratic centered $v_k$ with value $\phi_k$ with curvature $\gamma_k$.

        \par
        However, the construction exhibits variations for different variants of the accelerated gradient algorithm. 
        A different construction of $\phi_k$ yield a different algorithm. 
        As shown in Nesterov's book \cite*[(2.2.63)]{nesterov_lectures_2018}, a projected gradient variant of the accelerated gradient method make use of the following estimating functions, sequence: 
        \begin{align*}
            \phi_0 (x) &= f(x_0) + \frac{\gamma_0}{2}\Vert x - x_0\Vert^2
            \\
            l_f(x; y_k) &= f(\mathcal T_Ly_k) + 
            \frac{1}{2L}\Vert \mathcal G_L(y_k)\Vert^2 + 
            \langle \mathcal G_L(y_k), x - y_k\rangle + 
            \frac{\mu}{2}\Vert x - y_k\Vert^2
            \\
            \phi_{k + 1}(x) &= 
            (1 - \alpha_k)\phi_k + 
            \alpha_k l_f(x; y_k), 
        \end{align*}
        Here, $\mathcal G_L(x) = L(x - \mathcal T_L(x))$ and $T_L(x) = \Pi_Q(x - L^{-1}\nabla f(x))$, where $Q$ is a convex set. 
        It's equivalent to applying proximal gradient method with $h = f + g$ where $g = \delta_Q$. Finally, with $\mu = 0$, it reduces to ineuality appeared in 
        \hyperref[lemma:grad_map_linearization]{lemma \ref*{lemma:grad_map_linearization}}.

        \par
        Finally, the following constructions appeared in Guller \cite{guler_new_1992} are used to prove the convergence rate of accelerated PPM: 
        \begin{align*}
        \end{align*}
